---
title: "Covid91 vaccine study Final2022"
author: "Alex Moreno"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
Male = filter(FinalData, sex == "M")
Female = filter(FinalData, sex == "F")
LGBTQ  = filter(FinalData, LGBTQ == "gay")
DrugUser = filter(FinalData, DrugUser == "yes")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Hypothesis & Method

$H_0$: There is no difference between how effective the vaccine is for whichever sub-group uses it compared to the placebo. The values are independent

$H_A$: There is a difference between how effective the vaccine is for whichever sub-group uses it compared to the placebo. The values are dependent


Both variables are categorical. CAT~CAT is used for all different sub groups. Each group will have its own answer as to whether there is a difference or not. Each analysis will have a bar chart, chi-squared test and fisher exact test. Along with a brief summary for each sub group, the final conclusion will put together everything i evaluated and the hypothesis will be answered which is are the variables independent or dependent of each other. I will have similar hypothesis under each sub group to better anaylze my findings.


## Male

$H_0$: There is no relation between males & the vaccine.

$H_A$: There is a relation between males & the vaccine.


### Graphical Descriptive Results
```{r}
barchartGC(~infected + treatment, data = Male)
barchartGC(~infected + treatment, data = Male, type = "percent")
```

The frequency chart shows that those that took the placebo or vaccine had very little frequency of getting the virus. Those Males without either have gotten the virus at a much higher rate. The percentage chart also shows that those with the vaccine have a lesser chance to get the virus. Meanwhile the placebo while still shown as a low frequency , the virus is more stronger against the placebo than the vaccine in Males. 

### Numerical Descriptive Results
```{r}
table1 <- xtabs(~infected + treatment, data = Male)
rowPerc(table1)
colPerc(table1)
```

Those Males with the placebo can get more infected than those with the vaccine. The vaccine has a less percentage chance of getting the virus than the placebo. Those infected either have the placebo or nothing at all. The vaccine does work for Males. 

### Inferential Descriptive Results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The null hypothesis has to be rejected since the chi-squared test states that the p-value is 0.00000000004812. This p-value is less than 0.05. Chi-squared tests shows a statistic of 43.25 with degrees of freedom as 1. P-value for the chi-squared test is 0. Fisher exact test shows the p-value as 0.00000000002645. All the p-values are less than 0.05 so the null hypothesis is rejected in favor of the alternative. There is a relation between Males and the vaccine/placebo.


### Effictiveness and Confidence Intervals

```{r}
efficMale = 2.09/(2.09 + 1)*100
efficMale = round(efficMale)
confIntLower = 1.67/(1.67+1) * 100 
confIntLower = round(confIntLower)
confIntUpper = 2.64/(2.64 + 1 ) * 100 
confIntUpper = round(confIntUpper)
```

The lower confidence value is 63. The upper confidence value is 73. The avergae effectiveness for male is 68%. Since tha value is higher than 50% this proves the alternative hypothesis more. The vaccine gives a positive influence for Males.


### Summary Male

Males are dependent on the vaccine. The placebo had similar results to the vaccine but not exactly and not always. The vaccine for males is 68% effective. The null hypothesis is rejected since all the p-valus in the tests were below 0.05. So when a male takes the vaccine their chance of preventing the virus is greater than  those not taking it. Males should take the vaccine from these results.


## Female

$H_0$: There is no relation between females & the vaccine.

$H_A$: There is a relation between females & the vaccine.


### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment, data = Female)
barchartGC(~infected + treatment, data = Female, type = "percent")
```

From both graphs, females that take the vaccine have a lesser chance of getting the virus. Frequency graph shows those with the vaccine at a very low frequency for getting the virus. The percentage graph shows that those with the vaccine had a lesser chance than those without the vaccine or instead with the placebo. The vaccine seems to help Females.


### Numerical Descriptive Results
```{r}
table2 <- xtabs(~infected + treatment, data = Female)
rowPerc(table2)
colPerc(table2)
```
Females with the vaccine are less likely to get the virus. They have a 41% chance of getting the disease with the vaccine. The placebo has a 58.89% chance. And those without either have 49 and up percentages. The vaccine does give a better chance for females to prevent any contamination. 


### Inferential Descriptive Results 
```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
Chi-squared test has a p-value of 0.0011. This is less than 0.05 so the null hypothesis must be rejected in favor of the alternative. The chi-squared statistic is 10.653 while the degrees of freedom is 1. The fisher exact test also has a p-value that is less than 0.05 which is 0.0009484. Odds ratio is 1.4 which is not equal to 1 so the alternative hypothesis is true. 


### Effictiveness and Confidence Intervals

```{r}
efficFemale = 1.43/(1.43 + 1)*100
efficFemale = round(efficFemale)
confIntLower = 1.15/(1.15 + 1) * 100 
confIntLower = round(confIntLower)
confIntUpper = 1.78/(1.78 + 1 ) * 100 
confIntUpper = round(confIntUpper)
```

The efficacy value for Female is 59%. This value is higher than 50% so the vaccine is effective for females. If the percentage is higher than 50% it proves that the variable iss effective towards the other. The vaccine gives positive influence for Females.


### Summary Female

The vaccine does correlate with Female. The p-values from the analysis and test are all less than 0.05. The null hypothesis is rejected in favor of the alternative. Females that have the vaccine have a greater chance at preventing the virus. Females should take the vaccine from this analysis.


## LGBTQ

$H_0$: There is no relation between LGBTQ+ members & the vaccine.

$H_A$: There is a correlation between LGBTQ+ members & the vaccine.


### Graphical Descriptive Results 

```{r}
barchartGC(~infected + treatment, data = LGBTQ)
barchartGC(~infected + treatment, data = LGBTQ, type = "percent")
```


Those in this category that do take the vaccine seem to contract the virus more. The placebo is less likely to contract. The vaccine does not help the LGBTQ. The placebo helps more in this situation.

### Numerical Descriptive Results
```{r}
table3 <- xtabs(~infected + treatment, data = LGBTQ)
rowPerc(table3)
colPerc(table3)
```

Looking at the top table, those LGBTQ that take the vaccine have a 86% chance of getting the virus. While only 13% have a chance of getting it with the placebo. Those in this group are more likely to get infected if they have the vaccine.


### Inferential Descriptive Results
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

Chi-squared test has a p-value of 0.0000000132, a statistic of 32.302, and degrees of freedom of 1. P-value is less than 0.05. Fisher exact test has a p-value of 0.000000001642. This value is also less than 0.05. Alternative hypothesis is true since the odds ratio is 0.146 which is not equal to 1. Odds ratio has to 1 to 1 for the null to be true. In this case, being LGBTQ means that it does affect how the vaccine performs. 


### Effictiveness and Confidence Intervals

```{r}
efficLGBTQ = .15/(.15 + 1)*100
efficLGBTQ = round(efficLGBTQ)
confIntLower = .06/(.06 + 1) * 100 
confIntLower = round(confIntLower)
confIntUpper = .31/(.31 + 1 ) * 100 
confIntUpper = round(confIntUpper)
```

Efficacy value for LGBTQ is 13%. The vaccine is not effective well enough to prevent the virus for this subgroup. Since the value is less than 50, so the alternative hypothesis is accepted while the null is rejected. The vaccine gives a negative influence for the LGBTQ.


### Summary LGBTQ

From the analysis, those who identify as LGBTQ and take the vaccine do have a higher chance of getting infected still. The placebo seems to work better in this case. Being LGBTQ does affect the effectiveness of the vaccine so the null hypothesis is rejected for the alternative hypothesis. Since all p-values are below 0.05. The effectiveness for LGBTQ is not great. Those in this group should not take the vaccine and instead the placebo. 


## DrugUser

$H_0$: There is no relation between drug users & the vaccine.

$H_A$: There is a relation between drug users & the vaccine.

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment, data = DrugUser)
barchartGC(~infected + treatment, data = DrugUser, type= "percent")
```

Percentage graph shows the vaccine to be not effective for the DrugUser. The percentage shows that the placebo prevents it more than the actual vaccine. The vaccine has a higher percentage rate at around 85% for getting infected. A Drug User is more likely to get infected.


### Numerical Descriptive Results 
```{r}
table4 <- xtabs(~infected + treatment, data = DrugUser)
rowPerc(table4)
colPerc(table4)
```

The table shows how many get infected from the Druguser subgroup. 85% get infected with the vaccine. Less of these get infected with the placebo. The DrugUser should not take the vaccine based on the analysis so far. They are better off with the placebo. Seems as if Being a DrugUser does correlate to the vaccine. 


### Inferential Descriptive Results 
```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

Chi-squared test has a p=value of 0.0000001803. This value is less than 0.05 so the null hypothesis has to be rejected. The GC chi squared tests has a p-value of 0, also less than 0.05. Null hypothesis is rejected in favor of the alternative. Alternative hypothesis is also true due to the odds ratio not being equal to 1. Odds ratio is 0.1775. All p-values are below 0.05 and the fisher exact odds ratio is not equal to 1 so the null hypothesis is rejected. 


### Effictiveness and Confidence Intervals

```{r}
efficDrugUser = .18/(.18 + 1)*100
efficDrugUser = round(efficDrugUser)
confIntLower = .08/(.08 + 1) * 100 
confIntLower = round(confIntLower)
confIntUpper = .37/(.37 + 1 ) * 100 
confIntUpper = round(confIntUpper)
```

The efficacy of DrugUser is 15%. The vaccine is 15% succesful which in this case is not efficient. The value has to be greater than 50% for it to be statistically effective and efficient. The vaccine gives a negative influence for the DrugUser.

### Summary DrugUser

Does being a Drug User correlate to the vaccine? After the analysis this question can be answered as yes. If you are not a Drug User you will have a better chance at preventing the virus with the vaccine. The placebo helps out the Druguser more than the actual vaccine. Being a DrugUser is not good in this situation. DrugUsers should not take the vaccine from the results as they will most likely end up being infected still.


# Final Summary 

In this study, we investigated whether or not the preventative vaccine was effective for a particular subgroup. Male, Female, LGBTQ, or DrugUser. We analyzed if the placebo or vaccine worked better for each group individually. After analyzing the results, we can conclude that the vaccine does prevent only certain groups. So being a part of a certain group does affect how the vaccine works. Each group has its own analysis and all the groups reject the null hpothesis in favor of the alternative.
With Male, the p-values are all below 0.05 so the null hypothesis is rejected. The vaccine does help prevent the virus at a 68% efficacy rate. Vaccine is recommended.
With Female, the p-values are all below 0.05. The effectiveness rate is at 59% for Female. Vaccine is recommended. 
With LGBTQ, the p-values are all below 0.05. The placebo was more helpful in this group. Vaccine has an effective rate of 13% which is way below 50%. 
With DrugUser, the p-values were all below 0.05. The placebo also prevented the virus better than the actual vaccine. The efficacy rate is at 15%. The vaccine is not recommended.
The Male had the best results with the vaccine. The LGBTQ had the worst effective results. The alternative hypothesis is true because all values are different across each subgroup. For the null hypothesis to have been true, each group would have shown the same exact results no matter what. But that is not the case. Being a part of a group does correlate to the vaccines effectiveness in comparison to the placebo. Placebo worked better for the LGBTQ and DrugUser, while the vaccine worked better for the Male and Female. The vaccine is what was being tested thoroughly though. We wanted to know if the vaccine reacts different with these subgroups and the answer is yes it does react differently. 

